---
title: "R Notebook"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
#load libraries
library(llnl)
library(tidyverse)
```

Takes the abundance tables generated by coverm and creates normailized tables (reads per contig length per total metagenomic reads).

# Virome

```{r Viromes - Data}
mean.df <- read.csv("../raw_abundance_tables/virome/bulk_virome_0.75_0.9_trimmed_mean.csv",sep = '\t',row.names = 1)

total.df  <- read.csv("../raw_abundance_tables/virome/bulk_virome_total_reads.csv",sep = ',',row.names = 1) 

# Generate the normalized abundance table

bulk_virome_norm <- abundance.normal(mean.df,total.df,"total.bp" )
```

```{r}
#Remove votus with no matches
bulk_virome_norm.filtered <- bulk_virome_norm[rowSums(bulk_virome_norm)>0,]

write.csv(bulk_virome_norm.filtered,"../normalized_abundance_tables/bulk_virome_normalized_table.csv")

```

# Metatranscriptomes
```{r Data}

#Gene count data
gene.90 <- read.csv("../raw_abundance_tables/metat/metat.genes_0.9_counts.csv", sep = "\t",row.names = 1) %>% 
  filter(rowSums(across(where(is.numeric))) > 0) #remove genes from tables with no mapped reads, a little easier to deal with later on

# mean table for each votu
votu.90 <- read.csv("../raw_abundance_tables/metat/metat_0_0.9_mean.csv", row.names = 1, sep = "\t")
rownames(votu.90) <- make.names(rownames(votu.90), unique = TRUE) #This remove special characters that or originally in the row names

# mean table for each votu
votu.count.90 <- read.csv("../raw_abundance_tables/metat/metat_0.9_counts.csv", row.names = 1, sep = "\t")
colnames(votu.count.90) <- gsub("\\.Read\\.Count","",colnames(votu.count.90))
rownames(votu.count.90) <- make.names(rownames(votu.count.90), unique = TRUE) #This remove special characters that or originally in the row names

#Total metaT reads
total.df <- read.csv("../raw_abundance_tables/metat/metat_qcreads_stats.csv", row.names = 1)

# vOTU length
votu.length <- read.csv("../votu_length.tsv", sep = "\t")
```

Select for genes that are present

```{r}
#For the gene to be present, a gene must have at least 4 reads accross replicates, if not convert to 0

#Transpose the table
gene.t <- data.frame(t(gene.90))

#remove the End of the table
gene.t <- gene.t %>% 
  mutate(Samples = gsub(rownames(.), pattern = "\\.Read\\.Count", replacement = ""))

#Convert value to zero if the sum of the reads for each gene per sample (includes all replicates) is less than 4 and if it is greater than or equal to 4 it will leave as is

gene.active <- do.call(rbind, lapply(unique(gene.t$Samples), function(x) {
  df <- gene.t[gene.t$Samples == x, ]
  
  # Store the sample ID
  sample_id <- df$Samples[1]
  
  # Drop the Samples column for processing
  df_vals <- df %>% select(-Samples)
  
  # Get columns with sum < 4
  cols_low <- colSums(df_vals, na.rm = TRUE) < 4
  # Get columns with sum >= 4
  cols_high <- !cols_low
  
  # Set low-sum columns to 0
  df_vals[, cols_low] <- 0
  
  # Combine and add back the Samples column
  df_final <- cbind(Samples = sample_id, df_vals)
  
  return(df_final)
}))

rownames(gene.active) <- gene.active$Samples

gene.active <- gene.active %>% 
  select(-Samples)

#remove genes that do not have reads
gene.active.4.reads <- gene.active[,colSums(gene.active)>0]
```

Identify vOTUs with > 1 gene per 10kbp
```{r}
#Make a dataframe that has the votu identifier without the gene number

#remove rows if there sum is 0, removes genes with no reads
gene.active.t <- gene.active %>% 
  t() %>%
  as.data.frame() %>% 
  filter(rowSums(across(where(is.numeric))) > 0)%>%
  mutate(across(where(is.numeric), ~ ifelse(. > 0, 1, .))) %>% 
  mutate(gene_sum = rowSums(across(where(is.numeric))),
          vOTU = sub("_[^_]*$", "", rownames(.))) %>% 
  group_by(vOTU) %>%
  summarise(total_gene_sum = n(), .groups = "drop")
  # summarise(total_gene_sum = sum(gene_sum), .groups = "drop")

#Sub in the length name votus the - with a . to match the gene df
votu.length$X.name <- gsub("[-/|]", ".", votu.length$X.name)

#convert the length to kbp
votu.length$length.10kbp <- votu.length$length/10000

#Add vOTU length to the gene dataframe
gene.df.merged <- merge(gene.active.t,votu.length,by.x="vOTU",by.y="X.name")

#Calculate the number of genes per kbp of vOTU

gene.df.merged$gene.per.10kb <- gene.df.merged$total_gene_sum/gene.df.merged$length.10kbp

#filter for greater or equal to 1 gene per 10kbp
gene.df.filer <- gene.df.merged[gene.df.merged$gene.per.10kb>=1,]
no.gene.df.filer <- gene.df.merged[gene.df.merged$gene.per.10kb<1,]

```

Take votu table and filter it based on votus that are active

```{r}
#Filter for vOTUs that are active
votu.metat.abundance <- votu.90[rownames(votu.90) %in% gene.df.filer$vOTU,]

#Remove vOTUS that do not have reads
votu.metat.abundance <- votu.metat.abundance[rowSums(votu.metat.abundance) > 0, ]

#Drop the "mean" from the sample names
names(votu.metat.abundance) <- gsub("\\.Mean","",names(votu.metat.abundance))

#Filter the count table
votu.counts.filtered <- votu.count.90[rownames(votu.count.90) %in% rownames(votu.metat.abundance),]

#Frist check if column and rownames are the same order
same_rows <- identical(rownames(votu.counts.filtered), rownames(votu.metat.abundance))
same_cols <- identical(colnames(votu.counts.filtered), colnames(votu.metat.abundance))

votu.counts.filtered <- (votu.metat.abundance>0)*1 * votu.counts.filtered

votu.counts.sum <- data.frame(metat_mapped = colSums(votu.counts.filtered))

total.df <- cbind(total.df,
                  votu.counts.sum)
total.df <- total.df %>% 
  mutate(pct_metat_mapped = metat_mapped/total.reads *100)
```


Generate Normalized table
```{r}
votu.metat.ab.norm <- abundance.normal(votu.metat.abundance,total.df,"total.bp")

#Save table
write.csv(votu.metat.ab.norm,"../normalized_abundance_tables/active_votu_abundances.csv", row.names = TRUE)
write.csv(total.df,"../metat_read_metadata.csv")

```

# eDNA
```{r eDNA - Data}
edna.counts <- read.csv("../raw_abundance_tables/edna/edna_0.9_counts.csv",sep = '\t',row.names = 1)

rownames(edna.counts) <- make.names(rownames(edna.counts), unique = TRUE) #This remove special characters that or originally in the row names

colnames(edna.counts) <- gsub("\\.Read\\.Count", "", colnames(edna.counts)) # Remove the trailing sub-string that CoverM adds

edna.0 <- read.csv("../raw_abundance_tables/edna/edna_0_default_mean.csv",sep = '\t',row.names = 1)

rownames(edna.0) <- make.names(rownames(edna.0), unique = TRUE) #This remove special characters that or originally in the row names

colnames(edna.0) <- gsub("\\.Mean", "", colnames(edna.0)) # Remove the trailing sub-string that CoverM adds

total.df  <- read.table("../raw_abundance_tables/edna/total_reads_edna.txt",header=TRUE) %>% 
  mutate(sum_len = as.numeric(gsub(",", "", sum_len)),   # remove commas, convert to numeric
         num_seqs = as.numeric(gsub(",", "", num_seqs)),   # remove commas, convert to numeric
         SampleID = str_split(file,"_R",simplify = TRUE)[,1],
         SampleID = gsub("-","\\.",SampleID)) %>%  # Modifies sample IDs to make it easy to sum reads
  group_by(SampleID) %>%
  summarise(total.bp = sum(sum_len),
            total.seq = sum(num_seqs)) %>% # Sums the R1 and R2 reads to get the total reads
  column_to_rownames("SampleID")
```

Filter for votus in eDNA data that have >3 reads accross replicates
```{r}
# The functions to do this
filter_replicate_counts <- function(counts.df, min_reads = 3){
  filtered.df <- counts.df %>% 
    t() %>% 
    data.frame() %>% #Transposes the data frame so that we can create a grouping variable
    rownames_to_column(var = "SampleID") %>% 
    mutate(Group = str_sub(SampleID,1,nchar(SampleID)-2)) %>% #Remove the replicate ID
    group_by(Group) %>%
    mutate(across(where(is.numeric), ~ if (sum(.) > min_reads) . else 0)) %>% #filter for votus that have more than min_reads across replicates
    ungroup() %>% 
    column_to_rownames("SampleID") %>% 
    t() %>% 
    data.frame() #Convert back to columns being samples and rows being votus
}

#Apply the function to both breadth cutoffs
edna.0.counts.filtered <- filter_replicate_counts(edna.counts, min_reads = 3)


same_rows <- identical(rownames(edna.0), rownames(edna.0.counts.filtered))
edna.0.counts.filtered <- edna.0.counts.filtered[rownames(edna.0), colnames(edna.0)]%>% 
  mutate(across(everything(), ~ as.numeric(.)))
same_cols <- identical(colnames(edna.0), colnames(edna.0.counts.filtered))

edna.filtered.0 <- (edna.0.counts.filtered > 0)* 1 * edna.0
edna.0.normal <- abundance.normal(edna.filtered.0,total.df,"total.bp" )

```

Save normalized votu table
```{r}
#Remove votus with no matches
edna.0.normal.filtered <- edna.0.normal[rowSums(edna.0.normal)>0,]

write.csv(edna.0.normal.filtered,"../normalized_abundance_tables/edna_normalized_table.csv")


```

