---
title: "Untitled"
author: "GG"
date: "2024-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(llnl)
library(rstatix)
library(ggpubr)
library(tidyverse)
library(scales)
library(ggtext)
library(patchwork)
library(ggtext)
```

Generate figure 4 and supplementary figure 2

Used the equation found in Alexa paper

https://www.nature.com/articles/s41467-023-40835-4

```{r Import Data}
votu.metadata <- readRDS("../Data/votu_metadata.RDS")

metadata.sample.virome <- read_csv("../Data/metadata/bulk_metadata.csv") %>% 
  select("SampleID","Virome Dry  Weight (g)")

metadata.sample.metat <- read_csv("../Data/metadata/metat_metadata.csv") %>% 
  select("SampleID","Metat Dry  Weight (g)")

predicted.dna.weight <- read.csv("../Data/predicted.dna.weight.csv") %>% 
  mutate(Header = gsub(">","",Header),
         Header = gsub("\\|","\\.",Header),
         Header = gsub("-","\\.",Header),
         Header = gsub("/","_",Header))

#add the weight in kDa to the dataframe
votu.metadata <- merge(votu.metadata, predicted.dna.weight, by.x = "row.names", by.y = "Header") %>% 
  column_to_rownames("Row.names")

metat.rna.conc <- readRDS("../Data/metat.rna.conc.RDS")

dna.conc <- readRDS("../Data/jgi.final.dna.conc.RDS")

dna.conc <- dna.conc %>% 
  full_join(metat.rna.conc)

#Reads mapped
virome.mapped <- read.csv("../Data/bulk.virome.total.reads.csv",row.names = 1) 

meta.mapped <- read.csv("../Data/metat.rare.csv")
meta.mapped$sample <- gsub("-",".",meta.mapped$sample)

#Virome abundance tables
#Normalized abundances
virome.table <- read.csv("../Data/normalized_abundance_tables/bulk_virome_normalized_table.csv",row.names = 1)
#Raw counts with > 90% ANI
virome.counts <- read.csv("../Data/raw_abundance_tables/virome/bulk_virome_0.9_counts.csv",sep = "\t",row.names = 1)

#Data to calculate reads for ssDNA viruses
metat.table <- read.csv("../Data/normalized_abundance_tables/active_votu_abundances.csv",row.names = 1)
metat.counts <- read.csv("../Data/raw_abundance_tables/metat/metat_0.9_counts.csv",sep = "\t",row.names = 1)

# Modify special characters in row names to match the metaT abundance table row names
rownames(metat.counts) <- rownames(metat.counts) %>%
  gsub("\\|", ".", .) %>%
  gsub("/", ".", .) %>%
  gsub("-", ".", .)
```

```{r functions}
# Custom function for formatting to add scientific notation to the figures
format_superscript <- function(x) {
  ifelse(
    x == 0, 
    "0",  # Leave 0 as-is
    gsub("10\\^0?([1-9]+)", "10<sup>\\1</sup>", 
         gsub("e\\+?", " × 10^", scales::scientific(x))) # Format scientific notation
  )
}
```


#Strand Corrected Weight
Need to do a strand corrected weight for. The Da weight is calculated for a single stand of DNA, if it is dsDNA or unassigned going to multiple by 2
```{r}
#add the weight in kDa to the dataframe
votu.metadata <- votu.metadata %>% 
  mutate(strand.corr.da = ifelse(ssdna.dsdna != "ssDNA",
                                 Estimated.Weight..Da.*2,
                                 Estimated.Weight..Da.))
```

# Mapped reads
## Virome
Calculate the number of reads that were mapped to a vOTU (viral reads)
```{r}
#convert the virome table to +/- # This will make sure that the vOTU had a breath >= 75% 
virome.table[virome.table > 0] <- 1

#multiply the count table by the +/- table # if a vOTU was not present it will change the value to 0
virome.counts.pa <- virome.table*virome.counts

#Count the number of reads that mapped per sample
#convert to dataframe
total.virome.counts <- data.frame(virome.counts = colSums(virome.counts.pa))

# changes the row names to match the other dataframe
rownames(total.virome.counts) <- paste(rownames(total.virome.counts),"virome",sep = "")

#add to the length data
virome.mapped <- merge(virome.mapped,total.virome.counts, by.x = "row.names", by.y = "row.names")

```

# Get Virome [DNA]

```{r}
dna.conc.virome <- dna.conc %>%
  left_join(virome.mapped,by = c("Samples"= "Row.names")) %>% # convert mapped reads into a % (This is read mapped to a vOTU)
  mutate(mapped.pct = virome.counts/total.reads, # convert mapped reads into a % (This is read mapped to a vOTU)
         viral.dna.ng = dna.ng*mapped.pct) %>%  #then multiple the DNA by the mapped % to get viral DNA
  filter(Samples != "T3H218O_Avirome") #Remove this samples because there was not enough soil to extract/sequence

# Get the average length of vOTUs per Phosphorus groups and Non Phosphorus Groups 
avg.bp.virome <- rbind(
  votu.metadata %>% 
    select(virome_phosphorus,length,strand.corr.da) %>% #vOTUs found in phosphorus treatment 
    filter(virome_phosphorus == TRUE) %>% 
    summarize(avg.weight.da = mean(strand.corr.da)) %>% 
    mutate(sample.type = "Virome.Phosphorus"),
  votu.metadata %>% 
    select(virome_no.phosphorus,length,strand.corr.da) %>% #vOTUs found in no phosphorus treatment 
    filter(virome_no.phosphorus == TRUE) %>% 
    summarize(avg.weight.da = mean(strand.corr.da)) %>% 
    mutate(sample.type = "Virome.No.Phosphorus"),
  votu.metadata %>% 
    select(virome_T0.phosphorus,length,strand.corr.da) %>% # vOTUs found in dry soil
    filter(virome_T0.phosphorus == TRUE) %>% 
    summarize(avg.weight.da = mean(strand.corr.da)) %>% 
    mutate(sample.type = "Virome.Dry"))


#lets calculate virions per sample
dna.conc.virome <- dna.conc.virome %>% 
  mutate(SampleID = gsub("virome","",Samples)) %>% 
  left_join(metadata.sample.virome, by = "SampleID") %>% 
  mutate(Phosphorus = ifelse(
    grepl("P",Samples),
    "Phosphorus",
    "No Phosphorus"
  ),virions.da = ifelse(
    grepl("T0",Samples),
    (viral.dna.ng*6.0221E23)/((avg.bp.virome %>% filter(sample.type == "Virome.Dry"))[[1]]*1E9),
    ifelse(
      grepl("P",Samples),
    (viral.dna.ng*6.0221E23)/((avg.bp.virome %>% filter(sample.type == "Virome.Phosphorus"))[[1]]*1E9),
    (viral.dna.ng*6.0221E23)/((avg.bp.virome %>% filter(sample.type == "Virome.No.Phosphorus"))[[1]]*1E9)
  )
  ),
  virions.da.g = virions.da/`Virome Dry  Weight (g)`)

# add a group of T0 that will be phosphorus, to make sure they have the same starting point for dry soils
t0.dups <- dna.conc.virome %>% 
  filter(Time.Point == "T0") %>% 
  mutate(Phosphorus = "Phosphorus")

dna.conc.virome <- dna.conc.virome %>% 
  bind_rows(t0.dups)

#since we doubled the number T0 in the data frame I want to remove them for our stats
#this will make sure that there are no duplicates of T0 and convert the source to Dry for stats
dna.conc.virome.no.t0 <- dna.conc.virome %>% 
  dplyr::filter(Time.Point != "T0" )

dna.conc.virome.stat.df <- dna.conc.virome %>% 
  dplyr::filter(Time.Point == "T0" & Phosphorus == "Phosphorus") %>% 
  mutate(Phosphorus = "Dry") %>% 
  bind_rows(dna.conc.virome.no.t0)

#Use a linear model 
lm.mod <- lm(virions.da.g~Time.Point+Phosphorus,dna.conc.virome.stat.df)

shapiro.test(lm.mod$residuals)
aov.mod <- aov(lm.mod)
summary(aov.mod)
time.pho <- tukey_hsd(aov.mod,) %>% 
  mutate(Time.Point = ifelse(grepl("Phosphorus",term),
                             "Phosphorus",
                             "time"))


#now compare if treatment was significant at each time point
time.p <- dna.conc.virome.stat.df %>%
  filter(Time.Point!="T0") %>%
  group_by(Time.Point) %>%
  do({
    # Fit the model with aov to allow for Tukey's HSD test
    model <- lm(virions.da.g ~ Phosphorus, data = .)
    model <- aov(virions.da.g ~ Phosphorus, data = .)
    
    # Perform Tukey's HSD test
    tukey_results <- tukey_hsd(model)
  })  %>%
  ungroup()

all.stats <- rbind(time.p,time.pho) %>% 
  mutate(p.adj = p.adjust(p.adj,method = "BH")) %>% 
  add_significance("p.adj") %>% 
  mutate(xmin = case_when(
    Time.Point == "T1" ~ 2,
    Time.Point == "T2" ~ 3,
    Time.Point == "T3" ~ 4,
    group1 == "T0" ~ 1,
    group1 == "T1" ~ 2,
    group1 == "T2" ~ 3,
    group1 == "T3" ~ 4
  ),xmax = case_when(
    Time.Point == "T1" ~ 2,
    Time.Point == "T2" ~ 3,
    Time.Point == "T3" ~ 4,
    group2 == "T0" ~ 1,
    group2 == "T1" ~ 2,
    group2 == "T2" ~ 3,
    group2 == "T3" ~ 4
  ),
  y.position = ifelse(grepl("T0",Time.Point),max(dna.conc.virome$virions.da)/(10/1.07),max(dna.conc.virome$virions.da)/10)#divide by soil
  ) %>% 
  mutate(p.adj.signif = ifelse(term == "Phosphorus" ,gsub("\\*","‡",p.adj.signif),p.adj.signif),
         y.position = ifelse(term == "Phosphorus" ,max(dna.conc.virome$virions.da)/10,y.position))

virions.virome <- ggplot(dna.conc.virome,aes(x=Time.Point,y=virions.da.g, #divide by soil
                                             ,color = Phosphorus))+
  geom_smooth(aes(group = Phosphorus,fill = Phosphorus))+
  scale_color_manual(values= col.lnll)+
  scale_fill_manual(values= col.lnll)+
  stat_pvalue_manual(all.stats %>% filter(Time.Point == "T1" |
                                            Time.Point == "T2" |
                                            Time.Point == "T3"),
                     hide.ns = TRUE,
                     tip.length = 0,step.increase = 0)+ #filter for if there was a significant difference +/- phosphorus in each time point
  stat_pvalue_manual(all.stats %>% filter(Time.Point == "time"),
                     hide.ns = TRUE,
                     tip.length = 0,
                     step.increase = 0.05,size = 7)+ #filter for if there was a significant difference between time points
  theme_classic2()+
  xlab("Time (weeks)")+
  ylab("Virions per g of soil")+
  labs(color = "Treatment",
       fill  = "Treatment")+
  scale_y_continuous(
    labels = function(x) format_superscript(x)
  )+
  theme(axis.text.y = ggtext::element_markdown(),
        text = element_text(size = 20))+
    ## space for labels + allow drawing outside panel
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(t = 5, r = 5, b = 28, l = 5),
        axis.title.x = element_text(margin = margin(t = 38))) +

  ## Dry/Wet bar under the x-axis (factor positions: T0=1, T1=2, T2=3, T3=4)
  annotation_custom(
    grid::linesGrob(x = unit(c(0.0, 0.25), "npc"),
                    y = unit(c(-0.09, -0.09), "npc"),
                    gp = grid::gpar(col = "burlywood4", lwd = 4))
  ) +
  annotation_custom(
    grid::linesGrob(x = unit(c(0.25, 1.0), "npc"),
                    y = unit(c(-0.09, -0.09), "npc"),
                    gp = grid::gpar(col = "darkblue", lwd = 4))
  )+   # blue

  ## Labels for the bars
  annotate("text", x = 1.0, y = -Inf, label = "Dry",
           vjust = 3.7, size = 6, fontface = "bold", colour = "burlywood4") +
  annotate("text", x = 3, y = -Inf, label = "Wet",
           vjust = 3.7, size = 6, fontface = "bold", colour = "darkblue")
```

# Transcriptionally active
## Mapped reads
```{r}
#convert the active viruses table to +/- # This will make sure that the vOTU had a breath >= 75% 
metat.table[metat.table > 0] <- 1

#filter for votus in the active vOTU abundance table so that the df are the same size
metat.counts <- metat.counts[rownames(metat.counts) %in% rownames(metat.table),] 

#multiply the count table by the +/- table # if a vOTU was not present it will change the value to 0
metat.counts.pa <- metat.table*metat.counts

#Count the number of reads that mapped per sample
#convert to dataframe
total.metat.counts <- data.frame(metat.counts = colSums(metat.counts.pa))

#add to the vOTU length data
meta.mapped <- merge(meta.mapped,total.metat.counts, by.x = "sample", by.y = "row.names")

```


```{r}
rna.conc.metat <- metat.rna.conc %>%
  rename_with(~ gsub(" ", "\\.", .)) %>% 
  select(Samples,Total.metaT.RNA..ng..2) %>% #change samples names to match MetaT sample names
  mutate(Samples = gsub("virome","",Samples),
         Samples = gsub("T0_","T0\\.R\\.Dry\\.soil\\.",Samples),
         Samples = gsub("H2","\\.R\\.H2",Samples),
         Samples = gsub("P_","\\.P\\.",Samples),
         Samples = gsub("_","\\.",Samples),
         Time.Point = substr(Samples,1,2)) %>% 
  left_join(meta.mapped,by = c("Samples"="sample")) %>% # convert mapped reads into a % (This is read mapped to a vOTU)
  mutate(mapped.pct = metat.counts/total.reads, # convert mapped reads into a % (This is read mapped to a vOTU)
         metat.dna.ng = Total.metaT.RNA..ng..2*mapped.pct) #then multiple the DNA by the mapped % to get viral DNA

# Get the average length of vOTUs per Phosphorus groups and Non Phosphorus Groups 
avg.bp.metat <- rbind(
  votu.metadata %>% 
    select(metat_phosphorus,length,strand.corr.da) %>% 
    filter(metat_phosphorus == TRUE) %>% 
    summarize(avg.weight.da = mean(strand.corr.da)) %>% 
    mutate(sample.type = "Metat.Phosphorus"),
  votu.metadata %>% 
    select(metat_no.phosphorus,length,strand.corr.da) %>% # counting T0 as no phosphorus since it technically did not have phosphorus added
    filter(metat_no.phosphorus == TRUE) %>% 
    summarize(avg.weight.da = mean(strand.corr.da)) %>% 
    mutate(sample.type = "Metat.No.Phosphorus"),
  votu.metadata %>% 
    select(metat_T0.phosphorus,length,strand.corr.da) %>% # counting T0 as no phosphorus since it technically did not have phosphorus added
    filter(metat_T0.phosphorus == TRUE) %>% 
    summarize(avg.weight.da = mean(strand.corr.da)) %>% 
    mutate(sample.type = "Metat.T0"))

#lets calculate virocells per sample
rna.conc.metat <- rna.conc.metat %>% 
  left_join(metadata.sample.metat, by = c("Samples" = "SampleID")) %>% 
  mutate(Phosphorus = ifelse(
    grepl("P",Samples),
    "Phosphorus",
    "No Phosphorus"
  ),virions.da = ifelse(
    grepl("T0",Samples),
    (metat.dna.ng*6.0221E23)/((avg.bp.metat %>% filter(sample.type == "Metat.T0"))[[1]]*1E9),
  ifelse(
    grepl("P",Samples),
    (metat.dna.ng*6.0221E23)/((avg.bp.metat %>% filter(sample.type == "Metat.Phosphorus"))[[1]]*1E9),
    (metat.dna.ng*6.0221E23)/((avg.bp.metat %>% filter(sample.type == "Metat.No.Phosphorus"))[[1]]*1E9))))

#Since the metaT is transcriptionally active we need to convert into virocells. Since we do not know how many viruses are in a virocell we will look at multiple different virus/cell

list.numbers<- c(1,5,10,50,100,200)

add.virocells <- function(list.numbers,df){
  new.df <- do.call(cbind,lapply(list.numbers,function(FS){
    col.name.da <- paste("Virocell_da.",FS,sep = "")
    df[[col.name.da]] <- df$virions.da/FS
    df_ <- data.frame(df[[col.name.da]],df$Samples)
    names(df_) <- c(col.name.da, "Samples")
    return(df_)
    }))
  new.df <- new.df[, !duplicated(names(new.df))]#remove duplicate columns
  new.df <- merge(df,new.df,by = "Samples")
  return(new.df)
}

df.virocells <- add.virocells(list.numbers,rna.conc.metat)


# add a group of T0 that will be phosphorus
t0.dups <- df.virocells %>% 
  filter(Time.Point == "T0") %>% 
  mutate(Phosphorus = "Phosphorus")

df.virocells <- df.virocells %>% 
  bind_rows(t0.dups)

#Convert to a long dataframe to make into figures
df.virocells.long <- df.virocells %>% 
  pivot_longer(contains("Virocell"),values_to = "Virocells") %>% 
  mutate(virus.per.cell = str_extract(name, "(?<=_).*"),        
          Virocells.g = Virocells/`Metat Dry  Weight (g)`) #divide by soil



#since we doubled the number T0 in the data frame I want to remove them for our stats
df.virocells.long.stat.df <- df.virocells.long %>% 
  dplyr::filter(Time.Point != "T0" )

df.virocells.long.stat.df <- df.virocells.long %>% 
  dplyr::filter(Time.Point == "T0" & Phosphorus == "Phosphorus") %>% 
  mutate(Phosphorus = "Dry") %>% 
  bind_rows(df.virocells.long.stat.df)

lm.mod <- lm(Virocells.g~Time.Point+ Phosphorus,data = df.virocells.long.stat.df %>% 
                          filter(grepl("da",virus.per.cell)) %>% 
                                   mutate(virus.per.cell=gsub("da\\.","",virus.per.cell)) %>% 
                                            filter(virus.per.cell == 200))

shapiro.test(lm.mod$residuals)
aov.mod <- aov(lm.mod)
summary(aov.mod)
time.pho <- tukey_hsd(aov.mod) %>%
  mutate(Time.Point = ifelse(grepl("Phosphorus",term),
                             "Phosphorus",
                             "time"))

time.p <- df.virocells.long.stat.df %>% 
                          filter(grepl("da",virus.per.cell)) %>% 
                                   mutate(virus.per.cell=gsub("da\\.","",virus.per.cell)) %>% 
                                            filter(virus.per.cell == 200) %>%
  filter(Time.Point!="T0") %>%
  group_by(Time.Point) %>%
  do({
    # Fit the model with aov to allow for Tukey's HSD test
    model <- lm(Virocells.g~ Phosphorus, data = .)

    # Perform Tukey's HSD test
    tukey_results <- tukey_hsd(model)
  })  %>%
  ungroup()

all.stat <-rbind(time.p,time.pho) %>%  
  mutate(p.adj = p.adjust(p.adj,method = "BH")) %>% 
  mutate(xmin = case_when(
    Time.Point == "T1" ~ 2,
    Time.Point == "T2" ~ 3,
    Time.Point == "T3" ~ 4,
    group1 == "T0" ~ 1,
    group1 == "T1" ~ 2,
    group1 == "T2" ~ 3,
    group1 == "T3" ~ 4
  ),xmax = case_when(
   Time.Point == "T1" ~ 2,
    Time.Point == "T2" ~ 3,
    Time.Point == "T3" ~ 4,
    group2 == "T0" ~ 1,
    group2 == "T1" ~ 2,
    group2 == "T2" ~ 3,
    group2 == "T3" ~ 4
  ),
  y.position = ifelse(term == "Phosphorus",max(df.virocells.long$Virocells.g)/245,max(df.virocells.long$Virocells.g)/240*1.1)
  ) %>% 
  mutate(p.adj.signif = ifelse(term == "Phosphorus" ,gsub("\\*"," ‡",p.adj.signif),p.adj.signif),
         y.position = ifelse(term == "Phosphorus" ,max(df.virocells.long$Virocells.g)/180,y.position))

# find offset below x-axis
y_offset <- -0.08 * max(df.virocells.long %>% 
                          filter(grepl("da",virus.per.cell)) %>% 
                                   mutate(virus.per.cell=gsub("da\\.","",virus.per.cell)) %>% 
                                            filter(virus.per.cell == 200) %>% pull(Virocells.g), na.rm = TRUE)


virions.metat.200 <- ggplot(df.virocells.long %>% 
                          filter(grepl("da",virus.per.cell)) %>% 
                                   mutate(virus.per.cell=gsub("da\\.","",virus.per.cell)) %>% 
                                            filter(virus.per.cell == 200),aes(x=Time.Point,y=Virocells.g #gram of soil
                                                                              ,color = Phosphorus))+
  geom_smooth(aes(group = Phosphorus,fill = Phosphorus))+
  scale_color_manual(values= col.lnll)+
  scale_fill_manual(values= col.lnll)+
theme_classic2()+
  xlab("Time (weeks)")+
  ylab("Virocells per gram of soil")+
  labs(fill = "Treatment",
       color = "Treatment")+
  stat_pvalue_manual(all.stat %>% filter(Time.Point == "T1" |
                                            Time.Point == "T2" |
                                            Time.Point == "T3"),
                     hide.ns = TRUE,
                     tip.length = 0,step.increase = 0,size = 5)+ #filter for if there was a significant difference +/- phosphorus in each time point+
  stat_pvalue_manual(all.stat %>% filter(Time.Point == "time"),
                     hide.ns = TRUE,
                     tip.length = 0,
                     step.increase = 0.05,size = 7)+
  scale_y_continuous(
    labels = function(x) format_superscript(x)
  )+
  theme(axis.text.y = ggtext::element_markdown(),
        text = element_text(size = 20))+
    ## space for labels + allow drawing outside panel
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(t = 5, r = 5, b = 28, l = 5),
        axis.title.x = element_text(margin = margin(t = 38))) +

  ## Dry/Wet bar under the x-axis (factor positions: T0=1, T1=2, T2=3, T3=4)
  annotation_custom(
    grid::linesGrob(x = unit(c(0.0, 0.25), "npc"),
                    y = unit(c(-0.09, -0.09), "npc"),
                    gp = grid::gpar(col = "burlywood4", lwd = 4))
  ) +
  annotation_custom(
    grid::linesGrob(x = unit(c(0.25, 1.0), "npc"),
                    y = unit(c(-0.09, -0.09), "npc"),
                    gp = grid::gpar(col = "darkblue", lwd = 4))
  )+   # blue

  ## Labels for the bars
  annotate("text", x = 1.0, y = -Inf, label = "Dry",
           vjust = 3.7, size = 6, fontface = "bold", colour = "burlywood4") +
  annotate("text", x = 3, y = -Inf, label = "Wet",
           vjust = 3.7, size = 6, fontface = "bold", colour = "darkblue")


library(Cairo)

# Mean of virocells +/- phosphorus
df.virocells.long %>% 
  filter(grepl("da",virus.per.cell)) %>% 
  mutate(virus.per.cell=gsub("da\\.","",virus.per.cell)) %>% 
  filter(virus.per.cell == 200,
         Time.Point != "T0") %>% 
  group_by(Phosphorus) %>% 
  summarize(mean.virocells = mean(Virocells.g),
            sd.virocells = sd(Virocells.g))

# Mean of virocells time point
df.virocells.long %>% 
  filter(grepl("da",virus.per.cell)) %>% 
  mutate(virus.per.cell=gsub("da\\.","",virus.per.cell)) %>% 
  filter(virus.per.cell == 200,
         # Time.Point != "T0"
         ) %>% 
  group_by(Time.Point) %>% 
  summarize(mean.virocells = mean(Virocells.g),
            sd.virocells = sd(Virocells.g))
```


```{r}
library(patchwork)
virions.virocells <- virions.virome+virions.metat.200+plot_layout(axes = "collect",guides = "collect")+plot_annotation(tag_levels = 'A')

ggsave(plot =virions.virocells, "../figures/figure4.svg",height = 6, width = 12 )
ggsave(plot =virions.virocells, "../figures/figure4.png",height = 6, width = 12 )
```

# Supplamentary figure 2
```{r supp figures}

virions.metat <- ggplot(df.virocells.long %>% 
                          filter(grepl("da",virus.per.cell)) %>% 
                                   mutate(virus.per.cell=gsub("da\\.","",virus.per.cell)),aes(x=Time.Point,y=Virocells.g,color = Phosphorus))+
  geom_smooth(aes(group = Phosphorus,fill = Phosphorus))+
  facet_wrap(~virus.per.cell, scales = "free_y")+
  scale_color_manual(values= col.lnll)+
  scale_fill_manual(values= col.lnll)+
theme_classic2()+
  xlab("Time (Weeks)")+
  ylab("Virocells per Gram of Soil")+
  labs(fill = "Treatment",
       color = "Treatment")+
  scale_y_continuous(
    labels = function(x) format_superscript(x)
  )+
  theme(axis.text.y = ggtext::element_markdown())

labels.p <- df.virocells.long %>%
  filter(grepl("da", virus.per.cell), Phosphorus == "Phosphorus") %>%
  mutate(virus.per.cell = gsub("da\\.", "", virus.per.cell)) %>%
  group_by(virus.per.cell) %>%
  filter(Time.Point == max(Time.Point)) %>%
  summarise(
    x = max(Time.Point),
    y = median(log10(Virocells.g), na.rm = TRUE),
    .groups = "drop"
  )

p.metat <- ggplot(df.virocells.long %>% 
                          filter(grepl("da",virus.per.cell),
                                 Phosphorus == "Phosphorus") %>% 
                                   mutate(virus.per.cell=gsub("da\\.","",virus.per.cell)),aes(x=Time.Point,y=log10(Virocells.g),color = virus.per.cell))+
  geom_smooth(aes(group = virus.per.cell,fill = virus.per.cell))+
  facet_wrap(~Phosphorus, scales = "free_y")+
  scale_color_manual(values= c("1" = "#800F24",
                               "5" = "#931426",
                               "10" = "#A71930",
                               "50" = "#C04A5A",
                               "100" = "#D98A93",
                               "200" = "#EAB8BD"
                               ))+
  scale_fill_manual(values= c("1" = "#800F24",
                               "5" = "#931426",
                               "10" = "#A71930",
                               "50" = "#C04A5A",
                               "100" = "#D98A93",
                               "200" = "#EAB8BD"
                              ))+
theme_classic2()+
  xlab("Time (weeks)")+
  ylab("log10(Virocells per gram of soil)")+
  labs(fill = "Phosphorus (Viruses/Cell)",
       color = "Phosphorus (Viruses/Cell)")+
  geom_text(data = labels.p, 
            aes(x = x, y = y, label = virus.per.cell, color = virus.per.cell), 
            hjust = -0.1, 
            size = 4) + # Adjust `size` as needed
  theme(legend.position = "none") 

p.meta.all <- df.virocells.long %>% 
                          filter(grepl("da",virus.per.cell),
                                 Phosphorus == "Phosphorus") %>% 
                                   mutate(virus.per.cell=gsub("da\\.","",virus.per.cell))

labels.no.p <- df.virocells.long %>%
  filter(grepl("da", virus.per.cell), Phosphorus == "No Phosphorus") %>%
  mutate(virus.per.cell = gsub("da\\.", "", virus.per.cell)) %>%
  group_by(virus.per.cell) %>%
  filter(Time.Point == max(Time.Point)) %>%
  summarise(
    x = max(Time.Point),
    y = median(log10(Virocells.g), na.rm = TRUE),
    .groups = "drop"
  )

no.p.metat <- ggplot(df.virocells.long %>% 
                          filter(grepl("da",virus.per.cell),
                                 Phosphorus == "No Phosphorus") %>% 
                                   mutate(virus.per.cell=gsub("da\\.","",virus.per.cell)),aes(x=Time.Point,y=log10(Virocells.g),color = virus.per.cell))+
  geom_smooth(aes(group = virus.per.cell,fill = virus.per.cell))+
  facet_wrap(~Phosphorus, scales = "free_y")+
  scale_color_manual(values= c("1" = "gray30",
                               "5" = "gray40",
                               "10" = "gray50",
                               "50" = "gray60",
                               "100" = "gray70",
                               "200" = "gray80"
                               ))+
  scale_fill_manual(values= c("1" = "gray30",
                               "5" = "gray40",
                               "10" = "gray50",
                               "50" = "gray60",
                               "100" = "gray70",
                               "200" = "gray80"
                              ))+
theme_classic2()+
  xlab("Time (weeks)")+
  ylab("log10(Virocells per gram of soil)")+
  labs(fill = "No Phosphorus (Viruses/Cell)",
       color = "No Phosphorus (Viruses/Cell)")+
  geom_text(data = labels.no.p, 
            aes(x = x, y = y, label = virus.per.cell, color = virus.per.cell), 
            hjust = -0.1, 
            size = 4) + # Adjust `size` as needed
  theme(legend.position = "none") # Optional: remove legend if redundant

no.p.meta.all <- df.virocells.long %>% 
                          filter(grepl("da",virus.per.cell),
                                 Phosphorus == "Phosphorus") %>% 
                                   mutate(virus.per.cell=gsub("da\\.","",virus.per.cell))

#combine the figures
no.p.metat+p.metat+plot_layout(axes = "collect",guides = "collect")

ggsave("../figures/Supplamentaryfigure2.png")

## Stats ##

gam_mod.p <- mgcv::gam(log10(Virocells.g) ~ virus.per.cell * Time.Point,
                     method = "REML",
                     data=p.meta.all)


gam_mod.no.p <- mgcv::gam(log10(Virocells.g) ~ virus.per.cell * Time.Point,
                     method = "REML",
                     data=no.p.meta.all)

all.gam <- mgcv::gam(log10(Virocells.g) ~ virus.per.cell * Time.Point*trt,
                     method = "REML",
                     data = rbind(p.meta.all %>% 
                                    mutate(trt = "Phosphorus"),
                                  no.p.meta.all %>% 
                                    mutate(trt = "No Phosphorus")))

anova(gam_mod.p)
anova(gam_mod.no.p)
anova(all.gam)
```
